---
title: "DenverTravel"
author: "cyorir"
date: "September 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## City of Denver Travel Expenses

The city of Denver provides data about travel and procurement expenses by the city. These data are available here:

https://www.denvergov.org/opendata/dataset/city-and-county-of-denver-procurement-and-travel

At present (September 2018), data from 2012 to 2017 are provided. We can see tht the maximum expenditure was $3907.80 for 2012.

```{r travel_2012}
suppressPackageStartupMessages(library(openxlsx))
data_2012 <- read.xlsx("TravelCard_2012.xlsx")
summary(data_2012$Transaction.Amount)
```

We can pull in data from each year and see how the maximum expenditure has varied.

```{r travel_max}
data_2013 <- read.xlsx("TravelCard_2013.xlsx")
data_2014 <- read.xlsx("TravelCard_2014.xlsx")
data_2015 <- read.xlsx("TravelCard_2015.xlsx")
data_2016 <- read.xlsx("TravelCard_2016.xlsx")
data_2017 <- read.xlsx("TravelCard_2017.xlsx")
ta_max <- data.frame(year = c("2012"), max_expenditure = c(max(data_2012$Transaction.Amount)), stringsAsFactors = FALSE)
ta_max <- rbind(ta_max, c(year = "2013", max(data_2013$Transaction.Amount)))
ta_max <- rbind(ta_max, c(year = "2014", max(data_2014$Transaction.Amount)))
ta_max <- rbind(ta_max, c(year = "2015", max(data_2015$Transaction.Amount)))
ta_max <- rbind(ta_max, c(year = "2016", max(data_2016$Transaction.Amount)))
ta_max <- rbind(ta_max, c(year = "2017", max(data_2017$Transaction.Amount)))
ta_max
```

Let's combine the data from each year. In order to make use of the transaction dates, they must be converted to a standard format. The data for 2015 are particularly tricky, since that is when the dates switch from excel format to m/d/Y format. A similar method would be necessary to use the Payment.Date field.

```{r travel_combine}
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(janitor))
suppressPackageStartupMessages(library(dplyr))
data_cols <- colnames(data_2017)
colnames(data_2012) <- data_cols
colnames(data_2013) <- data_cols
colnames(data_2014) <- data_cols
colnames(data_2015) <- data_cols
data_2012_clean <- data_2012 %>% mutate(Transaction.Date = ymd(excel_numeric_to_date(as.numeric(Transaction.Date))))
data_2013_clean <- data_2013 %>% mutate(Transaction.Date = ymd(excel_numeric_to_date(as.numeric(Transaction.Date))))
data_2014_clean <- data_2014 %>% mutate(Transaction.Date = ymd(excel_numeric_to_date(as.numeric(Transaction.Date))))
data_2015_cleanA <- data_2015 %>% filter(!grepl("/",Transaction.Date)) %>% mutate(Transaction.Date = ymd(excel_numeric_to_date(as.numeric(Transaction.Date))))
data_2015_cleanB <- data_2015 %>% filter(grepl("/",Transaction.Date)) %>% mutate(Transaction.Date = mdy(Transaction.Date))
data_2015_clean <- rbind(data_2015_cleanA, data_2015_cleanB)
data_2016_clean <- data_2016 %>% mutate(Transaction.Date = mdy(Transaction.Date))
data_2017_clean <- data_2017 %>% mutate(Transaction.Date = mdy(Transaction.Date))
data_temp <- rbind(data_2012_clean,data_2013_clean)
data_temp2 <- rbind(data_2014_clean,data_2015_clean)
data_temp3 <- rbind(data_2016_clean,data_2017_clean)
data_temp <- rbind(data_temp, data_temp2)
data_all <- rbind(data_temp, data_temp3)
summary(data_all$Transaction.Date)
```

With the data combined, the transaction amount can be plotted over time.

```{r travel_plot}
suppressPackageStartupMessages(library(ggplot2))
ggplot(data = data_all, aes(x=Transaction.Date,y=Transaction.Amount)) +
  geom_point(alpha = 0.02)
```

We can check to see where the most money has been spent.

```{r travel_merchants}
totalSpent <- data_all %>% group_by(Merchant.Name) %>%
  summarise(TotalSpent = sum(Transaction.Amount)) %>%
  arrange(desc(TotalSpent))
head(totalSpent)
```

We can also check which which departments are spending the most on travel.

```{r travel_programs}
totalSpent <- data_all %>% group_by(Program.Area) %>%
  summarise(TotalSpent = sum(Transaction.Amount)) %>%
  arrange(desc(TotalSpent))
head(totalSpent)
```